\bf Model Visualization Techniques for a Social Network Model

### 2017-05-25 paper.Rnw
there's some code that's not running but it compiles now!

\begin{table}
\centering
\begin{tabular}{lll}
\textbf{Structural Effects} \\
outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ & \includegraphics[width=3cm]{img/outdegree}\\
reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji}$ & \includegraphics[width=3cm]{img/reciprocity}\\
transitive triplets  & $s_{i3}(x) = \sum_{j,h}& \includegraphics[width=3cm]{img/transtrip} x_{ij}x_{jh}x_{ih}$ \\
\textbf{Covariate Effects}\\
covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}z_j$ & \includegraphics[width=3cm]{img/covaralter} \\
covariate-ego  & $s_{i5}(x) = z_i\sum_j x_{ij}$ & \includegraphics[width=3cm]{img/covearego} \\
same covariate & $s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(z_i = z_j)$ & \includegraphics[width=3cm]{img/covarsim}%\\
%jumping transitive triplets  & $s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(z_i = z_h \neq z_j)$
\end{tabular}
\caption{\label{tab:effects} Some of the possible effects to be included in the stochastic actor-oriented models in RSiena. There are many more possible effects, but we only consider a select few here. For a complete list, see the RSiena manual \citep{RSiena}.}
\end{table}

@@

The \texttt{RSiena} software that fits the models to data also comes equipped with the \verbatim{sienaGOF()} function for examining model fit. This function ``assess[es] the fit of the model with respect to auxiliary statistics of networks" \citep[p.~53]{RSienamanual}. Examples of the auxiliary statistics are the out- or indegree distribution on the nodes, with the option for users to input their own statistics to examine. The fit is tested using Mahalanobis distance to determine a $p$-value when comparing the statistics calculated from the observed data (excluding the first observation that is conditioned on) to the distribution of the statistics when calculated on the simulated waves. So, good fit is acheived when $p$ is large. XX more detail about mahalanobis distance and calculating p-value XX

The \verbatim{plot.sienaGOF()} function allows us to visualize this fit. Below is an example of what the goodness-of-fit plot output looks like for indegree and outdegree statistics for a small model. Box plots and violin plots are drawn on top of each other in Figure~\ref{fig:sienagof} in order to show the distribution of the simulated network observations compared to the true data shown in red points connected by red lines. If the red points lie well within the simulated values, the model is a good fit to the data. The model examined in Figure~\ref{fig:sienagof} appears to do a better job of capturing the outdegree distribution than the indegree distribution of the data.

@@

We also explore simulations from these different models given parameter values. Using the means from the 1,000 fitted parameter values, we simulated 1,000 observations from each of our three models. We condition on the first friendship network observation, and the second and third observations are simulated from the SAOM models estimated previously. From these simulations, we first create a visualization that represents an ``average" network. To do this, we count occurences of each possible edge in the simulations, resulting in a summary network with weighted edges representing the number of times an edge appeared in the simulated wave 2 when simulating from the SAOM 1000 times. In Figure~\ref{averagenet}, we show the first wave, the ``average'' network from the three models we fit, and the second wave. Comparing the three averages to waves 1 and 2, we see that they have very similar structure to wave 2. Model 2, which included the transitive triplet parameter, seems to have created more group strucutre overall than models 1 and 3. This could be an indication that, despite the highly significant nature of the parameter $\beta_3$ according to the $t$-tests in \verbatim{RSiena}, the transitive triplet parameter should not be included because it does not result in a structure similar to the data.

>>>>

\begin{table}
\centering
\begin{tabular}{lll}
\textbf{Structural Effects} \\
outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ & \includegraphics[width=2cm]{img/outdegree}\\
reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji}$ & \includegraphics[width=2cm]{img/reciprocity}\\
transitive triplets  & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ & \includegraphics[width=1.5cm]{img/transtrip}  \\
\textbf{Covariate Effects}\\
covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}z_j$ & \includegraphics[width=3cm]{img/covaralter} \\
covariate-ego  & $s_{i5}(x) = z_i\sum_j x_{ij}$ & \includegraphics[width=3cm]{img/covarego} \\
same covariate & $s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(z_i = z_j)$ & \includegraphics[width=3cm]{img/covarsim}%\\
%jumping transitive triplets  & $s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(z_i = z_h \neq z_j)$
\end{tabular}
\caption{\label{tab:effects} Some of the possible effects to be included in the stochastic actor-oriented models in RSiena. There are many more possible effects, but we only consider a select few here. For a complete list, see the RSiena manual \citep{RSiena}.}
\end{table}

@@

The \texttt{RSiena} software that fits the models to data also comes equipped with the \texttt{sienaGOF()} function for examining model fit. This function ``assess[es] the fit of the model with respect to auxiliary statistics of networks" \citep[p.~53]{RSienamanual}. Examples of the auxiliary statistics are the out- or indegree distribution on the nodes, with the option for users to input their own statistics to examine. The fit is tested using Mahalanobis distance to determine a $p$-value when comparing the statistics calculated from the observed data (excluding the first observation that is conditioned on) to the distribution of the statistics when calculated on the simulated waves. So, good fit is acheived when $p$ is large. XX more detail about mahalanobis distance and calculating p-value XX

The \texttt{plot.sienaGOF()} function allows us to visualize this fit. Below is an example of what the goodness-of-fit plot output looks like for indegree and outdegree statistics for a small model. Box plots and violin plots are drawn on top of each other in Figure~\ref{fig:sienagof} in order to show the distribution of the simulated network observations compared to the true data shown in red points connected by red lines. If the red points lie well within the simulated values, the model is a good fit to the data. The model examined in Figure~\ref{fig:sienagof} appears to do a better job of capturing the outdegree distribution than the indegree distribution of the data.

@@

We also explore simulations from these different models given parameter values. Using the means from the 1,000 fitted parameter values, we simulated 1,000 observations from each of our three models. We condition on the first friendship network observation, and the second and third observations are simulated from the SAOM models estimated previously. From these simulations, we first create a visualization that represents an ``average" network. To do this, we count occurences of each possible edge in the simulations, resulting in a summary network with weighted edges representing the number of times an edge appeared in the simulated wave 2 when simulating from the SAOM 1000 times. In Figure~\ref{averagenet}, we show the first wave, the ``average'' network from the three models we fit, and the second wave. Comparing the three averages to waves 1 and 2, we see that they have very similar structure to wave 2. Model 2, which included the transitive triplet parameter, seems to have created more group strucutre overall than models 1 and 3. This could be an indication that, despite the highly significant nature of the parameter $\beta_3$ according to the $t$-tests in \texttt{RSiena}, the transitive triplet parameter should not be included because it does not result in a structure similar to the data.
